### 2023-06-30 paper/paper.tex
add paper template

\documentclass[12pt]{article}

\include{./wits_template.zex}

\author{gg}
\title{test}
\begin{document}
    \maketitle
    gg
\end{document}